---
title: "Lecture 8: Covariance and Correlation"
output: 
  learnr::tutorial:
    css: css/styles.css
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
if (!require(kableExtra)) install.packages("kableExtra")
library(learnr)
library(tidyverse)
library(ggplot2)
library(kableExtra)
```


## Example: Discrete Random Variables


(Adopted from Exercise 10.3 from MIPS)

Let $U$ and $V$ be two random variables with joint probability distribution defined by the following probability mass function.


```{r echo=FALSE}
jpmf <- rbind(
  c("1/4", "0", "1/4", "1/2"),
  c("0", "1/2", "0", "1/2"),
  c("1/4", "1/2", "1/4", "1")
  )
rownames(jpmf) <- c("0", "1", "P(U=a)")
colnames(jpmf) <- c("0", "1", "2", "P(V=b)")
kable(jpmf) %>%
  kable_paper() %>%
  pack_rows("b", 1, 2) %>%
  pack_rows("", 3, 3) %>%
  add_header_above(c(" " = 1, "a" = 3, " " = 1))
```


Compute $E[UV]$ and $E[U+V]$ via 100,000 simulations.

$E[UV]$

```{r ex103, exercise=TRUE, exercise.lines=10}
N <- 10^5


```

```{r ex103-hint-1}
N <- 10^5
U <- sample(0:2, N, replace = TRUE,
            prob = c(.25, .5, .25))
V <- ifelse(U == 1, 1, 0)

```

```{r ex103-solution}
N <- 10^5
U <- sample(0:2, N, replace = TRUE,
            prob = c(.25, .5, .25))
V <- ifelse(U == 1, 1, 0)
sum(U * V) / N
sum(U + V) / N
```


## Weekly Activity 4: Q4 b)

The following code creates a data frame representing
a deck of playing cards. The data frame is named `deck` and is available in the exercise code chunk.

```{r echo=TRUE}
suits <- c("Spades", "Hearts", "Clubs", "Diamonds")
numbers <- c("A", 2:10, "J", "Q", "K")
values <- c(11, 2:10, 10, 10, 10)
deck <- data.frame(
  suit = rep(suits, each = 13),
  number = rep(numbers, 4),
  value = rep(values, 4)
)
```

```{r draw2-setup}
suits <- c("Spades", "Hearts", "Clubs", "Diamonds")
numbers <- c("A", 2:10, "J", "Q", "K")
values <- c(11, 2:10, 10, 10, 10)
deck <- data.frame(
  suit = rep(suits, each = 13),
  number = rep(numbers, 4),
  value = rep(values, 4)
)
```


Generate 1,000 simulates of the following game played by one person.

_Game step:_

i. From a randomly shuffled single deck, you draw 2 cards without replacement.
ii. Sum the values of the cards based on the counting rule described below. If the count is less than $17$ go to step iii. Otherwise, __end the game__.
iii. Draw an extra card from the remaining deck. 

  
_Counting rule:_

+ Counts are based on the `number`s.
+ Values between `2` and `10` are always counted as is.
+ Consider the face cards where the `number` is `J`, `Q`, or `K` as $10$.
+ Consider `A` as $11$ if the resulting total is less than $22$. Otherwise, consider `A` as $1$.
+ The "default" values are stored in `value` column of the data frame `deck`. `A` stored as $11$.

Let $S$ be the sum of the values you have on your hands at the end of the game. Draw a histogram of the simulated values of $S$ with `binwidth = 1` and estimate $P(S>21)$, $E[S]$, and $\text{Var}(S)$.

Today, we will simulate only the first pair of draws.

```{r draw2, exercise=TRUE, exercise.lines=15}
deck
```


```{r draw2-solution}
N <- 1000
S <- numeric(N)
for (i in seq_len(N)) {
  shuffled_index <- sample.int(nrow(deck))
  n_drawn <- 2
  s_tmp <- sum(deck$value[shuffled_index][1:n_drawn])
  if (s_tmp > 21) {
    s_tmp <- s_tmp - 10
  }
  S[i] <- s_tmp
}
ggplot() +
  theme_classic() +
  geom_histogram(aes(x = S), binwidth = 1)
```
