---
title: "Lecture 8: Covariance and Correlation"
output: 
  learnr::tutorial:
    css: css/styles.css
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggplot2)
library(kableExtra)
```


## Example: Discrete Random Variables


(Adopted from Exercise 10.3 from MIPS)

Let $U$ and $V$ be two random variables with joint probability distribution defined by the following probability mass function.


```{r echo=FALSE}
jpmf <- rbind(
  c("1/4", "0", "1/4", "1/2"),
  c("0", "1/2", "0", "1/2"),
  c("1/4", "1/2", "1/4", "1")
  )
rownames(jpmf) <- c("0", "1", "P(U=a)")
colnames(jpmf) <- c("0", "1", "2", "P(V=b)")
kable(jpmf) %>%
  kable_paper() %>%
  pack_rows("b", 1, 2) %>%
  pack_rows("", 3, 3) %>%
  add_header_above(c(" " = 1, "a" = 3, " " = 1))
```


Compute $E[UV]$ and $E[U+V]$ via simulations.

### $E[UV]$

```{r euv, exercise=TRUE, exercise.lines=10}
n <- 100000
```


### $E[U+V]$

```{r eupv, exercise=TRUE, exercise.lines=10}
n <- 100000
```

## Example: Correlation Coefficient

(Exercise 10.20 from MIPS)

Determine $\rho\left(U,U^2\right)$ when $U\sim\text{U}(0,a)$ for $a=2$ and $a=200$. In class, we computed the value to be $\sqrt{15}/2$= `r sqrt(15)/2` for any $a>0$


```{r echo=FALSE}
u <- runif(1000)
ggplot() +
  theme_classic() +
  geom_point(aes(x = u, y = u^2), size = 1, alpha = 0.05) +
  labs(x = "U", y = expression(U^2),
       title = expression(paste("1,000 simulated values of U and ", U^2))) +
  scale_x_continuous(breaks = c(0, 1), limits = c(0, 1), labels = c("0", "a")) +
  scale_y_continuous(breaks = c(0, 1), limits = c(0, 1), labels = c("0", "a")) +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 18)
  )
```

```{r rhouu, exercise=TRUE, exercise.lines=20}
# Write your R codes here.
n <- 100000
a <- 2

```


```{r rhouu2, exercise=TRUE, exercise.lines=20}
# Write your R codes here.
n <- 100000
a <- 200

```




## Example: Correlation and Independence

Let $X\sim N(0,1)$ and $Y=X^2$. Are they correlated?


```{r echo=FALSE}
x <- rnorm(1000)
ggplot() +
  theme_classic() +
  geom_point(aes(x = x, y = x^2), size = 1, alpha = 0.05) +
  labs(x = "X", y = "Y", title = "1,000 simulated values of U and Y") +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 18)
  )
```


```{r rhonorm, exercise=TRUE, exercise.lines=10}
# Write your R codes here.
n <- 100000

```

